---
title: Descriptive Statistics for Tidy Tuesday Alt Text Project
author: "Silvia P. Canelón, Elizabeth Hare"
date: February 24, 2021
output: html_document
---

# Summary: How many TidyTuesday posts have alt text?


``` {r, include=FALSE}
### don't display chunks in output doc by default
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(results='hide')
knitr::opts_chunk$set(fig_path=".")
library(knitr)
library(gtsummary)
library(TidyTuesdayAltText)
library(qdap) # word_count, character_count

### Access data from the four years
data(package="TidyTuesdayAltText")
data(ttTweets2018, ttTweets2019, ttTweets2020, ttTweets2021)
### concatenate year data.frames into one
ttTweets4y <- rbind(ttTweets2018, ttTweets2019, ttTweets2020, ttTweets2021)
table(ttTweets4y$Year)

### create factor variable describing alt text based on the four categories
### 1. Customized alt text written by author
### 2, Field contains default text: "Image"
### 3. Scraping error - Tweet is unavailable (deleted or private)
### 4. NA Unavailable due to non-static media


ttTweets4y$AltTextCat <- ifelse(is.na(ttTweets4y$AltText), "UnavailableNonStatic",
                         ifelse(ttTweets4y$AltText == "Image", "DefaultDescription",
                         ifelse(ttTweets4y$AltText == "ScrapingError", "ScrapingError",
                                "DescriptiveText")))
table(ttTweets4y$AltTextCat, useNA="ifany")
### convert AltTextCat to factor for easier analysis
ttTweets4y$AltTextCat <- as.factor(ttTweets4y$AltTextCat)


### make df with only image descriptions = "alt text tweets"
att <- ttTweets4y[ttTweets4y$AltTextCat == "DescriptiveText", ]
dim(att)

### cound words using qdap package
### qdap::word_count, parameters digit.remove=FALSE because we expect numerals
att$wordCount <- word_count(att$AltText, digit.remove=FALSE)
summary(att$wordCount)

att$charCount <- character_count(att$AltText)
summary(att$charCount)
```

<!-- Silvia: Minor change here to include a more specific date range -->

Silvia scraped `r nrow(ttTweets4y)` TidyTuesday Tweets posted between `r format(min(ttTweets4y$TweetDate), "%B %d, %Y")` and `r format(max(ttTweets4y$TweetDate), "%B %d, %Y")`. I concatenated them into one data set.

``` {r, results="asasis"}
tbl_cross(ttTweets4y, row=Year, col=AltTextCat,
          percent="row")
```


# Word and character counts

## Words

Words were counted using the `word_count()` function in the `qdap` R package, with
the option to include digits set to `TRUE` because numbers have important meaning
in describing graphs. The median word count for alt text descriptions was 
`r median(att$wordCount)`, with a range of 
`r min(att$wordCount)` to 
`r max(att$wordCount)`. 
The following table presents these statistics by year.


``` {r, results="asis"}
# minor changes here perhaps due to changes in gtsummary. Silvia downloaded gtsummary v1.3.7

tbl_summary(att, by=Year,
            include=c(Year, wordCount),
            type=all_continuous() ~ "continuous",
            statistic=all_continuous() ~ c("{median}, {min}-{max}"))

```

## Characters

<!-- Silvia: I'm not sure what was meant by the first line of the paragraph. As an aside, the maximum number of characters allowed for Twitter media alt-text is 1,000-->

The maximum character count for Twitter alt-text is 420. The median character count for all years is `r median(att$charCount)`, with a range of `r range(att$charCount)[1]` to `r range(att$charCount)[2]`. The following table provides these statistics by year:

``` {r, results="asis"}

tbl_summary(att, by=Year,
            include=c(Year, charCount),
            type=all_continuous() ~ "continuous",
            statistic=all_continuous() ~ c("{median}, {min}-{max}"))

```

# How many TidyTuesday posts include an external link?

``` {r, results="asasis"}
tbl_cross(ttTweets4y, row=Year, col=UrlCheck,
          percent="row")
```
