---
title: Descriptive Statistics for Tidy Tuesday Alt Text Project
author: "Silvia P. Canelón, Elizabeth Hare"
date: 
output: html_document
---


# Exploring Alt-Text Contents

``` {r, setup, include=FALSE, echo=FALSE}
library(knitr)
library(TidyTuesdayAltText)
library(gtsummary)

### what data sets are in updated package?
data(package="TidyTuesdayAltText")
data(altTextSubset)
### load Silvia's annotated file
annotatedS <- readRDS("../data/AltTextSubset_Annotated.rds")
colnames(annotatedS)
### add NA columns for rubric because you can't add them in Numbers
annotatedS$conveysMeaning <- NA
annotatedS$axisInfo <- NA
annotatedS$axisOrValueScale <- NA
annotatedS$plotType <- NA
annotatedS$repeatTitle <- NA

annotatedL <- annotatedS[ ,c("TweetId", "AltText", "conveysMeaning", "axisInfo",
                             "axisOrValueScale", "plotType", "repeatTitle",
                             "TitleText", "SubtitleText",
                             "OtherText", "AnnotationNotes")]
### keep only rows with alt-texts
### how many missing
dim(annotatedL[is.na(annotatedL$AltText), ])
### remove "scraping error" in altTExt field
annotatedL2 <- annotatedL[!annotatedL$AltText == "ScrapingError", ]
dim(annotatedL2)
head(annotatedL2[ , 1:3])
tail(annotatedL2[ ,1:3])
```

``` {r, write csv for blank rubric, eval=FALSE}
### write file to enter rubric data
write.csv(annotatedL2, file="annotated.csv", row.names=FALSE)

```

I started with the file "data/AltTextSubset_Annotated.rdsg5
" from Silvia. I kept only the columns I needed, 
excluding the hashtag variable that contained lists that couldn't be written easily to .csv.
I also removed rows where AltText == "ScrapingError". I ended upwith 
`r nrow(annotatedL2)` observations.  

I added columns for the rubric because I was having trouble adding rows to the spreadsheet where.
I evaluated each alt-text on five items:

- ```conveysMeaning``` Can I understand something about what the data is showing from the alt-text (e.g.
the relationship between variables for a line graph or scatter plot)?
* ```axisInfo``` Does the alt-text tell me what the axes are? NA if the graphic is something
like a map or pie chart
- ```axisOrValueScale``` Does the alt-text contain information on the ranges of the axes, or 
describe the magnitude of the findings in another way (e.g. individual bar heights on bar charts)?
- ```plotType``` Does the alt-text tell me what kind of graphic this is (line graph, bar chart, 
scatterplot, map, pie chart)?
- ```repeatTitle``` Is the alt-text a repetition of the title or subtitle?  

This is a subjective evaluation in that it is made without any access to the graphic, 
and I'm sure I'm using my knowledge of different kinds of graphs to fill in some of the gaps.
Also I'm not familiar with some of the newer, fancier types of plots (waterfall?).  

I saved this as a .csv and filled the rubric item fields in Numbers. I saved my results as a .csv
and read it back into R.



``` {r, read in csv with rubric items filled in, include=FALSE, echo=FALSE}
annotatedRubric <- read.csv("annotatedRubric1.csv")
### remove if conveys meaning == NA
### NA if it's a repeat, or I think it is a picture without data
annotatedRubric <- annotatedRubric[!is.na(annotatedRubric$conveysMeaning), ]
dim(annotatedRubric)
table(annotatedRubric$conveysMeaning, useNA="ifany")
table(annotatedRubric$axisInfo, useNA="ifany")
table(annotatedRubric$axisOrValueScale, useNA="ifany")
### 9 is probably typo for 0
annotatedRubric$axisOrValueScale[annotatedRubric$axisOrValueScale == 9] <- 0
table(annotatedRubric$plotType, useNA="ifany")
table(annotatedRubric$repeatTitle, useNA="ifany")
### 9 is typo for 0
annotatedRubric$repeatTitle[annotatedRubric$repeatTitle == 9] <- 0

```

After removing duplicate tweets and tweets that seemed to contain only photos or screenshots that 
didn't contain data visualizations, there were `r nrow(annotatedRubric)` tweets that contained alt-text. The following table contains frequencies of presence (1) or absence (0) of alt-text rubric 
items.

``` {r, frequency table for rubric items, results="asis", echo=FALSE}
rubricTab <- tbl_summary(annotatedRubric, 
                         include=c(conveysMeaning:repeatTitle),
                         type=everything() ~ "categorical",
                         statistic=all_categorical() ~ "{n} ({p})",
                         missing="always",
                         missing_text="NA")
rubricTab



```
